---
title: "Trail for preprocessing - Survival Analysis"
output: html_notebook
---

# Imports

```{r}
library(readr)
```

# Data reading

```{r}
hatchery <- read_csv("../raw_data/hatchery.csv", show_col_types = FALSE)

downstream <- read_csv("../raw_data/downstream.csv", show_col_types = FALSE)

microtroll_hatch <- read_csv("../raw_data/microtroll_hatch.csv", 
                             show_col_types = FALSE)

microtroll_wild <- read_csv("../raw_data/microtroll_wild.csv", 
                            show_col_types = FALSE)
```

The columns we want the final dataset to have is the following:

```{r}
columns <- colnames(microtroll_hatch)
columns
```

### Note on `action` column

`Action` can be taken care only after combining the whole dataset. The first instance of a tag_id is "tag" and the following instances is "detect".

# Pre-processing

## Stage 1: Hatchery

Notes: Table clean, good to go

```{r}
head(hatchery)
```

## Stage 2: Downstream

Notes: Fix date column, rest good to go

```{r}
head(downstream)
```

## Stage 3: Estuary

```{r}

```

## Stage 4: Microtroll

Notes: These data sets are in the required format. No further processing needed.

Action column - what to do?

```{r}
head(microtroll_hatch)
```

```{r}
head(microtroll_wild)
```

## Stage 5: Return

```{r}

```

# Combining cleaned dfs

```{r}

```

## `Action` column 

```{r}

```

# Save the dataset

```{r}

```
